---
title: "Language"
date: 2020-05-02 16:39:22
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 2
bibliography: a_problem_course_in_mathematical_logic.bib
link-citations: yes
---

<style>

body {
text-align: justify}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = TRUE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}

library(tidyverse)
library(latex2exp)
```

# Language

This section is based on [@bilaniuk_problem_2009, Chapter 1]

The primary objective of propositional logic, $\mathcal{L}_P$, is to formalize the reasoning that can be done using connectives like _not_, _and_, _or_ and _if ... then_

__Definition__ The symbols of $\mathcal{L}_P$ are

- Parentheses: $($ and $)$
- Connectives: $\neg$ and $\to$
- Atomic formulas: $A_0, A_1, A_2, \dots , A_n, \dots$

Now we specify how the symbols of $\mathcal{L}_P$ can be put together

__Definition__ The formulas of $\mathcal{L}_P$ are any finite sequences of $\mathcal{L}_P$ symbols which satisfy the following rules

- Every atomic formula is a formula
- If $\alpha$ is a formula then $(\neg \alpha)$ is a formula
- If $\alpha$ and $\beta$ are formulas then $\alpha \to \beta$ is a formula
- No other sequence of symbols in $\mathcal{L}_P$ are formulas

    + The parentheses are used to group other symbols together
    + The atomic formulas represent statements that cannot be broken down any further using $\neg$ or $\rightarrow$
    + Every atomic formula is a formula and every other formula is built from shorter formulas using $\neg$, $\rightarrow$, $($ and $)$ in particular ways
    + The last part points out that there are exactly three rules to form formulas in $\mathcal{L}_P$
    
## Problems 1-7 and solutions
    
__Problem 1__ _Why  are the  following not formulas of $\mathcal{L}_P$? There might be more than one reason_ $\dots$

- $A_{-56}$

    + $A_{ \color{red}{-56}}$ is not an atomic formula
    
- $(Y \to A)$

    + $\color{red}{Y}$ and $\color{red}{A}$ are not atomic formulas
    
- $(A_7 \leftarrow A_4)$

    + $\color{red}{\leftarrow}$ is not a connective

- $A_7 \to (\neg A_5))$

    + $($ is missing at the beginning of the sequence
    
        + $\color{red}{(}A_7 \to (\neg A_5))$
    
- $(A_8A_9 \to A_1043998$
    
    + $)$ is missing at the end of the sequence
    
        + $(A_8A_9 \to A_1043998 \color{red}{)}$
    
    + $\color{red}{A_8A_9}$ is not a formula
    
- $(((\lnot A_1) \to (A_\ell \to A_7) \to A_7)$
    
    + One $)$ is missing or $($ and $)$ are missing 
    
        + $(((\lnot A_1) \to (A_\ell \to A_7) \color{red}{)} \to A_7)$
        
        + $(((\lnot A_1) \to \color{red}{(} (A_\ell \to A_7) \to A_7) \color{red}{)}$
            
    + $A_{\color{red}{\ell}}$ is not an atomic formula
    
__Problem 2__ _Show that every formula of $\mathcal{L}_P$ has the same number of left  parentheses as it has of right parentheses_

- Proof by strong induction where $n$ is the number of connectives in a formula of $\mathcal{L}_P$

    + _Base step_ $(n = 0)$: If $\gamma$ is a formula with 0 connectives it must be an atomic formula. Therefore $\gamma$ has the same number of $($ and $)$.
    
    + _Induction hypothesis_ $(n \leq k)$: If $\gamma$ is a formula with $n \leq k$ connectives then it has the same number of $($ and $)$.
    
    + _Induction step_ $(n = k + 1)$: If $\gamma$ is a formula with $n = k + 1$ connectives then we have two possibilities:
    
        + $\gamma = (\neg \alpha)$ where $\alpha$ has $n = k$ connectives
        
        + $\gamma = (\alpha \to \beta)$ where $\alpha$ and $\beta$ have $n \leq k$ connectives
        
            + If $\gamma = (\neg \alpha)$ then by the _induction hypothesis_ $\alpha$ has the same number of $($ and $)$. Therefore $\gamma$ also has the same number of $($ and $)$
            
            + If $\gamma = (\alpha \to \beta)$ then by the _induction hypothesis_ $\alpha$ and $\beta$ has the same number of $($ and $)$. Therefore $\gamma$ also has the same number of $($ and $)$
            
__Problem 3__ _Suppose $\alpha$ is any formula of $\mathcal{L}_P$.  Let $\ell(\alpha)$ be the length of $\alpha$ as a sequence of symbols and let $p(\alpha)$ be the number of parentheses (counting both left and right parentheses) in $\alpha$.  What are the minimum and maximum values of $p(\alpha) / \ell(\alpha)$?_

- The minimum values of $p(\alpha)$ and $\ell(\alpha)$ are 0 and 1 which is the case when $\alpha$ is an atomic formula. Therefore the minimum value of $p(\alpha) / \ell(\alpha)$ is 0 

- To find the maximum value of $p(\alpha) / \ell(\alpha)$ it is important to point out that:

    + $\ell(\alpha) = p(\alpha) + c(\alpha) + a(\alpha)$ where $c(\alpha)$ and $a(\alpha)$ are the number of connectives and atomic formulas in $\alpha$
    
    + $p(\alpha) = 2c(\alpha)$
    
    + To maximize $p(\alpha) / \ell(\alpha)$ we need to minimize $\ell(\alpha)$ by choosing the minimum value of $a(\alpha)$
    
        + The minimum value of $a(\alpha)$ is 1 and corresponds to formulas built from negating a formula at each step and starting with an atomic formula: $A_0, (\neg A_0), (\neg (\neg A_0)), \dots$
        
        + For this case we have that $p(\alpha) / \ell(\alpha) = \frac{2c(\alpha)}{3c(\alpha) + 1}$ where $c(\alpha) \in \{ 0, 1, 2, \dots, n\}$
        
            + $p(\alpha) / \ell(\alpha)$ doesn't have a maximum but when $c(\alpha) \to \infty$ then $p(\alpha) / \ell(\alpha) \to 2 /3 \approx 0.667$ 
            
```{r}

ggplot(tibble(x = 0:100), aes(x)) + 
    stat_function(fun   = ~ 2 * .x / (3 * .x + 1) , 
                  geom  = "point",
                  shape = 21,
                  fill  = "#E31A1C") + 
    geom_hline(yintercept = 2/3,
               color      = "#FF7F00") +
    scale_y_continuous(breaks = seq(0, 4/6, 1/6), 
                       labels = scales::number_format(accuracy = 0.001)) +
    labs(x = TeX("$c(\\alpha)$"),
         y = TeX("$\\frac{2c(\\alpha)}{3c(\\alpha) + 1}$"))  +
    theme(panel.border      = element_rect(fill = NA, color = "black"),
          plot.background   = element_rect(fill = "#f3fcfc"),          
          panel.background  = element_rect(fill = "#f3f7fc"))
```

__Problem 4__ *Suppose $\alpha$ is any formula of $\mathcal{L}_P$.  Let $a(\alpha)$ be the number of atomic formulas in $\alpha$ (counting repetitions) and let $c_{\to}(\alpha)$ be the number of occurrences of $\to$ in $\alpha$. Show that $a(\alpha) = c_{\to}(\alpha) + 1$.*

- Proof by strong induction where $n$ is th number of connectives in a formula  

    + _Base step_ $(n = 1)$: If $\gamma$ is a formula with 0 connectives it must be an atomic formula. Therefore $\gamma$ has 0 occurences of $\to$ and 1 occurence of an atomic formula.
    
    + _Induction hypothesis_ $(n \leq k)$: If $\gamma$ is a formula with $n \leq k$ connectives then $a(\gamma) = c_{\to}(\gamma) + 1$ 
    
    + _Induction step_ $(n = k + 1)$: If $\gamma$ is a formula with $n = k + 1$ connectives then we have two possibilities:
    
        + $\gamma = (\neg \alpha)$ where $\alpha$ is a formula with $n = k$ connectives
        
        + $\gamma = (\alpha \to \beta)$ where $\alpha$ and $\beta$ have $n \leq k$ connectives
        
            + If $\gamma = (\neg \alpha)$ then by the _induction hypothesis_ $a(\alpha) = c_{\to}(\alpha) + 1$. Therefore for $\gamma$ we have that $a(\gamma) = c_{\to}(\gamma) + 1$
            
            + If $\gamma = (\alpha \to \beta)$ then by the _induction hypothesis_ $a(\alpha) = c_{\to}(\alpha) + 1$ and $a(\beta) = c_{\to}(\beta) + 1$. Therefore:
            
                + $c_{\to}(\gamma) = c_{\to}(\alpha) + c_{\to}(\beta) + 1$
            
                + $\begin{split}
                   a(\gamma) & = a(\alpha) + a(\beta) \\
                   & = c_{\to}(\alpha) + 1 + c_{\to}(\beta) + 1\\
                   & = c_{\to}(\alpha) + c_{\to}(\beta) + 1 + 1 \\
                   & = c_{\to}(\gamma) + 1
                   \end{split}$

__Problem 5__ _What are the possible lengths of formulas of $\mathcal{L}_P$? Prove it._

We have that 

- $\ell(\alpha) = p(\alpha) + c(\alpha) + a(\alpha)$

- $p(\alpha) = 2c(\alpha)$

- $c(\alpha) = c_{\neg}(\alpha) + c_{\to}(\alpha)$ where $c_{\neg}(\alpha)$ is the number of occurrences of $\neg$ in $\alpha$

- $a(\alpha) = c_{\to}(\alpha) + 1$

    + Therefore $\ell(\alpha) = 4c_{\to}(\alpha) + 3c_{\neg}(\alpha)$ where $c_{\to}(\alpha) \in \{ 0, 1, 2, \dots, n\}$ and $c_{\neg}(\alpha) \in \{ 0, 1, 2, \dots, n\}$
    
        + We can fix $c_{\neg}(\alpha) = 0, 1, 2, 3$ in the rows and use $c_{\to}(\alpha) \in \{ 0, 1, 2, \dots n\}$ in the columns to build the following matrix:
        
            + $\begin{matrix}
                              & \color{red}{0} & \color{red}{1} & \color{red}{2} & \color{red}{3} & \cdots & \color{red}{n}  \\
               \color{red}{0} &              1 &              5 &              9 &             13 & \cdots & 4n + 1          \\
               \color{red}{1} &              4 &              8 &             12 &             16 & \cdots & 4n + 4          \\
               \color{red}{2} &              7 &             11 &             15 &             19 & \cdots & 4n + 7          \\
               \color{red}{3} &             10 &             14 &             18 &             22 & \cdots & 4n + 10         \\
               \end{matrix}$
            
Therefore we can see that the possible lengths of formulas of $\mathcal{L}_P$ are $1, 4, 5$ and $7, 8, 9, 10, \dots, n$

__Problem 6__ _Find a way for doing without parentheses or other punctuation symbols in defining a formal language for propositional logic._

A possible way is to use the _polish notation_ invented by Jan Lukasiewicz as mentioned in [@church_introduction_1956, p 38]. Another option is to use _reverse polish notation_ used in some Hewlett-Packard calculators

- In our case _polish notation_ is apply in the following way:

    + $(\alpha \to \beta)$ can be written as $\to \alpha \beta$

    + $(\neg \alpha)$ can be written as $\neg \alpha$

        + For example $((\neg \alpha) \to \beta)$, $(\alpha \to (\neg \beta))$, $((\neg \alpha) \to (\neg \beta))$, $(\neg (\alpha \to \beta))$, $(\neg ( (\neg \alpha) \to \beta))$, $(\neg ( \alpha \to (\neg \beta)))$ and $(\neg ( (\neg \alpha) \to (\neg \beta)))$ are written as:
    
            + $\to \neg \alpha \beta$
        
            + $\to \alpha \neg \beta$
        
            + $\to \neg \alpha \neg \beta$
        
            + $\neg \to \alpha \beta$
        
            + $\neg \to \neg \alpha \beta$
        
            + $\neg \to \alpha \neg \beta$
        
            + $\neg \to \neg \alpha \neg \beta$
            
- In our case _reverse polish notation_ is apply in the following way:

    + $(\alpha \to \beta)$ can be written as $\alpha \beta \to$

    + $(\neg \alpha)$ can be written as $\alpha \neg$

        + For example $((\neg \alpha) \to \beta)$, $(\alpha \to (\neg \beta))$, $((\neg \alpha) \to (\neg \beta))$, $(\neg (\alpha \to \beta))$, $(\neg ( (\neg \alpha) \to \beta))$, $(\neg ( \alpha \to (\neg \beta)))$ and $(\neg ( (\neg \alpha) \to (\neg \beta)))$ are written as:
    
            + $\alpha \neg \beta \to$
        
            + $\alpha \beta \neg \to$
        
            + $\alpha \neg \beta \neg \to$
        
            + $\alpha \beta \to \neg$
        
            + $\alpha \neg \beta \to \neg$
        
            + $\alpha \beta \neg \to \neg$
        
            + $\alpha \neg \beta \neg \to \neg$
            
__Problem 7__ _Show that the set of formulas of $\mathcal{L}_P$ is countable._

Let's define the following sets

- $S_{\mathcal{L}_P} = \{ (, ), \neg, \to, A_0, A_1, A_2, \dots, A_n, \dots \}$ the set of symbols of $\mathcal{L}_P$

    + Because $\{ A_0, A_1, A_2, \dots, A_n, \dots \}$ is countable then $S_{\mathcal{L}_P}$ is countable
    
        + $S_{\mathcal{L}_P}^i = \underbrace{S_{\mathcal{L}_P} \times \dots \times S_{\mathcal{L}_P}}_\text{i times}$ for $i = 1, 4, 5, 7, 8, 9, 10, \dots, n$ are countable 

            + $S_{\mathcal{L}_P} = \bigcup_{i = 1}^n S_{\mathcal{L}_P}^i$ is also countable
          
-  $F_{\mathcal{L}_P \ell_i}$ the formulas of $\mathcal{L}_P$ separated by its individual components in a _i-tuple_ where $i = 1, 4, 5, 7, 8, 9, 10, \dots, n$. Each $F_{\mathcal{L}_P \ell_i}$ represents the formulas of $\mathcal{L}_P$ with length $i$.

    + For example $(A_0 \to (\neg A_1))$ can be represented as $[ (, A_0, \to, (, \neg, A_1, ), ) ]$ where $[ (, A_0, \to, (, \neg, A_1, ), ) ] \in F_{\mathcal{L}_P \ell_8}$ 

        + $F_{\mathcal{L}_P} = \bigcup_{i = 1}^n F_{\mathcal{L}_P \ell_i}$
    
            + $F_{\mathcal{L}_P}$ is countable because $F_{\mathcal{L}_P} \subset S_{\mathcal{L}_P}$
        
- Therefore the set of formulas of $\mathcal{L}_P$ is countable

## Informal Conventions I

We will use the symbols $\land$,  $\lor$,  and $\leftrightarrow$ \index to represent _and_,  _or_ and _if and only if_. Since they are not among the symbols of $\mathcal{L}_P$,  we will use them as abbreviations for certain constructions involving only $\lnot$ and $\to$:

- $(\alpha \land \beta)$ is short for $(\neg (\alpha \to (\neg \beta)))$ 
- $(\alpha \lor \beta)$ is short for $((\neg \alpha) \to \beta)$
- $(\alpha \leftrightarrow \beta)$ is short for $((\alpha \to \beta) \land (\beta \to \alpha))$

## Problems 8-9 and solutions

__Problem 8__ _Take a couple of english sentences with several connectives and translate them into formulas of $\mathcal{L}_P$.  You may use $\land$,  $\lor$,  and $\leftrightarrow$ if appropriate._

- I will go to Peru or Argentina only if I am on vacations

    + $A_0$: I will go to Peru
    
    + $A_1$: I will go to Argentina
    
    + $A_2$: I am on vacations
    
        + $((A_0 \lor A_1) \leftrightarrow A_2)$
        
- If neither I nor you can do it then anyone can't do it

    + $A_0$: I can do it
    
    + $A_1$: You can do it
    
    + $A_3$: Anyone can do it
    
        + $((\neg (A_0 \lor A_1)) \to (\neg A_3))$
        
__Problem 9__ _Write out $((\alpha \lor \beta) \land (\beta \to \alpha))$ using only $\neg$ and $\to$._

- $(\neg (((\neg \alpha) \to \beta) \to (\lnot (\beta \to \alpha))))$

## Informal Conventions II

For the sake of readability, we will occasionally use some informal conventions that let us get away with writing fewer parentheses:

- We will usually drop the outermost parentheses in a formula

    + $\alpha \to \beta$ instead of $(\alpha \to \beta)$
    
- We will let $\neg$ take precedence over $\to$ when parentheses are missing

    + $\neg \alpha \to \beta$ instead of $((\neg \alpha) \to \beta)$
    
        + Also fit the informal connectives into this scheme by letting the order of precedence be $\neg, \land, \lor, \to$ and $\leftrightarrow$
        
- We will group repetitions of $\to, \lor, \land, \leftrightarrow$ to the right when parentheses are missing

    + $\alpha \to \beta \to \gamma$ instead of $(\alpha \to (\beta \to \gamma))$
    
## Problems 10-11 and solutions

__Problem 10__ Write out $\neg (\alpha \leftrightarrow \neg \delta ) \land \beta \to \neg \alpha \to \gamma$ first with the missing parentheses included and then as an official formula of $\mathcal{L}_P$.

- With parentheses

    + $(((\neg (\alpha \leftrightarrow (\neg \delta))) \land \beta) \to ((\neg \alpha) \to \gamma))$
    
- Using only $\mathcal{L}_P$ symbols

    + $((\neg ((\neg (\neg ((\alpha \to (\neg \delta)) \to (\neg ((\neg \delta) \to \alpha))))) \to (\neg \beta))) \to ((\neg \alpha) \to \gamma))$

__Definition__ Suppose $\varphi$ is a formula of $\mathcal{L}_P$.  The set of _subformulas_ of $\varphi$,  $\mathcal{S}(\varphi)$,  is defined as follows:

- If $\varphi$ is an atomic formula,  then $\mathcal{S}(\varphi) = \{ \varphi \}$

- If $\varphi$ is $(\neg \alpha)$,  then $\mathcal{S}(\varphi) = \mathcal{S}(\alpha) \cup \{ (\neg \alpha) \}$.

- If $\varphi$ is $(\alpha \to \beta)$,  then $\mathcal{S}(\varphi) = \mathcal{S}(\alpha) \cup \mathcal{S}(\beta) \cup \{ (\alpha \to \beta) \}$

For example

- $A_4 \to A_1 \lor A_4$

    + This is equivalent to $(A_4 \to (A_1 \lor A_4))$
    
        + This is equivalent to $(A_4 \to ((\neg A_1) \to A_4))$
        
            + $\mathcal{S}((A_4 \to ((\neg A_1) \to A_4))) = \mathcal{S}(A_4) \cup \mathcal{S}(((\neg A_1) \to A_4)) \cup \{ (A_4 \to ((\neg A_1) \to A_4)) \}$
            
                + $\mathcal{S}(((\neg A_1) \to A_4)) = \mathcal{S}((\neg A_1)) \cup \mathcal{S}(A_4) \cup \{ ((\neg A_1) \to A_4) \}$
                
                    + $\mathcal{S}((\neg A_1)) = \{ (\neg A_1) \}$

__Problem 11__ _Find all the subformulas of each of the following formulas_

- $(\neg ((\neg A_{56}) \to A_{56}))$

    + $\mathcal{S}((\neg ((\neg A_{56}) \to A_{56}))) = \{ (\neg ((\neg A_{56}) \to A_{56})), ((\neg A_{56}) \to A_{56}), (\neg A_{56}), A_{56} \}$ 
    
- $A_9 \to A_8 \to \neg (A_{78} \to \neg \neg A_0)$

    + $A_9 \to A_8 \to \neg (A_{78} \to \neg \neg A_0)$ is equivalent to $(A_9 \to (A_8 \to (\neg (A_{78} \to (\neg (\neg A_0))))))$
    
        + $\begin{split}
           \mathcal{S}(A_9 \to A_8 \to \neg (A_{78} \to \neg \neg A_0)) =  
           & \{ (A_9 \to (A_8 \to (\neg (A_{78} \to (\neg (\neg A_0)))))) \} \\ 
           & \cup \{ (A_8 \to (\neg (A_{78} \to (\neg (\neg A_0))))) \} \\
           & \cup \{ (\neg (A_{78} \to (\neg (\neg A_0)))) \} \\
           & \cup \{ (A_{78} \to (\neg (\neg A_0))) \} \\ 
           & \cup \{ (\neg (\neg A_0)) \} \\ 
           & \cup \{ (\neg (\neg A_0)) \} \\ 
           & \cup \{ (\neg A_0) \} \\ 
           & \cup \{ A_{78}, A_9, A_8, A_0 \} \\
           \end{split}$
    
- $\neg A_0 \land \neg A_1 \leftrightarrow \neg (A_0 \lor A_1)$

    + $\neg A_0 \land \neg A_1 \leftrightarrow \neg (A_0 \lor A_1)$ is equivalent to $(((\neg A_0) \land (\neg A_1)) \leftrightarrow (\neg (A_0 \lor A_1)))$
    
        + $\begin{split}
           \mathcal{S}(\neg A_0 \land \neg A_1 \leftrightarrow \neg (A_0 \lor A_1)) =
           & \{ (((\neg A_0) \land (\neg A_1)) \leftrightarrow (\neg (A_0 \lor A_1))) \} \\
           & \cup \{ ((\neg A_0) \land (\neg A_1) \} \\
           & \cup \{ (\neg (A_0 \lor A_1))) \} \\
           & \cup \{ (A_0 \lor A_1) \} \\
           & \cup \{ (\neg A_0), (\neg A_1) \} \\
           & \cup \{ A_1, A_0 \}
           \end{split}$
           
## Unique readability

The formulas of $\mathcal{L}_P$ can only be read in exactly one way if all the symbols of $\mathcal{L}_P$ are distinct and no symbol is a subsequence of any other symbol. With this additional assumption, one can prove the following:

__Theorem__ If $\gamma$ is a formula of $\mathcal{L}_P$ it must satisfy exactly one of the following conditions:

- $\gamma$ is an atomic formula and is unique
- There is a unique formula $\alpha$ such that $\gamma = (\neg \alpha)$
- There are unique formulas $\alpha$ and $\beta$ such that $\gamma = (\alpha \to \beta)$

__Proof__

- $\gamma$ must satisfy at least one of the conditions because $\gamma$ is a formula of $\mathcal{L}_P$
    
- $\gamma$ must satisfy exactly one condition because:
    
    + If $\gamma$ is an atomic formula then $\gamma$ doesn't contain parentheses so
          $\gamma \neq (\neg \alpha)$ and $\gamma \neq (\alpha \to \beta)$
          
    + If $\gamma = (\neg \alpha)$ then assume that $\gamma = (\varphi \to \beta)$. Therefore
        
        + $\begin{split}
           \gamma = (\neg \alpha) \land \gamma = (\varphi \to \alpha) & \iff \neg \alpha) = \varphi \to \beta) \\
           & \implies \varphi \text{ begins with } \neg \\
           & \implies \text{There is a contradiction because } \varphi \text{ is a formula}  
              \end{split}$
              
        + Therefore it is not the case that $\gamma = (\neg \alpha) \text{ and } \gamma = (\varphi \to \alpha)$
            
    + If $\gamma$ satisfy exactly one of those conditions then:
        
        + If $\gamma$ is an atomic formula then it is unique because all the symbols of $\mathcal{L}_P$ are distinct
            
        + If $\gamma = (\neg \alpha)$ and $\gamma = (\neg \varphi)$ then:
            
            + $\begin{split}
                   \gamma = (\neg \alpha) \text{ and } \gamma = (\neg \varphi) & \iff \neg \alpha) = \neg \varphi) \\
                   & \iff \neg \alpha = \neg \varphi \\
                   & \iff \alpha = \varphi
                   \end{split}$
                   
        + If $\gamma = (\alpha \to \beta)$ and $\gamma = (\varphi \to \theta)$ we need to prove that if $\sigma$ and $\tau$ are formulas of $\mathcal{L}_p$ with $\sigma \neq \tau$ it is not the case that $\sigma$ begins with $\tau$, to show that $\alpha = \beta \land \varphi = \theta$ [See [@enderton_mathematical_2001, p 30, Lemma 13B] for details]. Using this result we have that:
        
        
            + $\begin{split}
               \gamma = (\alpha \to \beta) \text{ and } \gamma = (\varphi \to \theta) & \iff \alpha \to \beta) = \varphi \to \theta) \\
               & \iff \alpha \to \beta = \varphi \to \theta \\
               & \iff \alpha = \varphi \text{ and } \to \beta = \to \theta \\
               & \iff \alpha = \varphi \text{ and } \beta = \theta
                   \end{split}$
        
To understand the importance of __unique readability__ check out [@zach_sets_2019, Chapter 5, Section 5.4] 

# Bibliography
