---
title: "Truth Assignments"
date: 2020-05-19 20:29:29
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 2
bibliography: a_problem_course_in_mathematical_logic.bib
link-citations: yes
---

<style>

body {
text-align: justify}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = TRUE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}


```

# Truth Assignments

This section is based on [@bilaniuk_problem_2009, Chapter 2]

A formula $\varphi$ of $\mathcal{L}_P$ is true or false depending on how we interpret the atomic formulas which appear in $\varphi$.

- Since we don't want to commit ourselves to a single interpretation we will define how any assignment of _truth values_ $T$ ("true") and $F$ ("false") to atomic formulas of $\mathcal{L}_P$ can be extended to all other formulas.

We will also get a reasonable definition of what it means for a formula of $\mathcal{L}_P$ to follow logically from other formulas. 

__Definition__ A _truth assignment_ is a function $v$ whose domain is the set of all formulas of $\mathcal{L}_P$ and whose range is the set $\{ T, F \}$ of truth values,  such that:

- $v(A_n)$ is defined for every atomic formula $A_n$

- For any formula $\alpha$

    + $v((\neg \alpha)) = \begin{cases}
                           T & \text{if } v(\alpha) = F \\
                           F & \text{if } v(\alpha) = T
                           \end{cases}$

- For any formulas $\alpha$ and $\beta$

    + $v((\alpha \to \beta)) = \begin{cases}
                                F & \text{if } v(\alpha) = T \text{ and } v(\beta) = F \\
                                T & \text{Otherwise}
                                \end{cases}$
                                
Note that we have not defined how to handle any truth values besides $T$ and $F$ in $\mathcal{L}_P$. According to [@bilaniuk_problem_2009, Chapter 2] logics with other truth values have uses,  but are not relevant in most of mathematics.


A convenient way to write out the determination of the truth value of a formula on a given truth assignment is to use a _truth table_: 

- List all the subformulas of the given formula across the top in order of length

- Fill in their truth values on the bottom from left to right

    + Except for the atomic formulas at the extreme left,  the truth value of each subformula will depend on the truth values of the subformulas to its left
    
        + For example if $v(A_0) = T$ and $v(A_1) = F$ then to know the truth assignment of $v(((\neg A_1) \to (A_0 \to A_1)))$ we can use the following truth table:
        
$$\begin{array}{|c|c|c|c|c|}
   A_0 & A_1 & (\neg A_1) & (A_0 \to A_1) & ((\neg A_1) \to (A_0 \to A_1)) \\
   \hline
   T   & F   & T          & F             & F
   \end{array}$$

## Problems 1-10 and solutions

__Problem 1__ Suppose $v$ is a truth assignment such that $v(A_0) = v(A_2) = T$ and $v(A_1) = v(A_3) = F$.  Find $v(\alpha)$ if $\alpha$ is:

- $\neg A_2 \to \neg A_3$

$$\begin{array}{|c|c|c|c|c|}
   A_2 & A_3 & (\neg A_2) & (\neg A_3) & ((\neg A_2) \to (\neg A_3)) \\
   \hline
   T   & F   & F          & T          & T
   \end{array}$$

- $\neg A_2 \to A_3$

$$\begin{array}{|c|c|c|c}
   A_2 & A_3 & (\neg A_2) & ((\neg A_2) \to A_3) \\
   \hline
   T   & F   & F          & T
   \end{array}$$

- $\neg ( \neg A_0 \to A_1)$

$$\begin{array}{|c|c|c|c|c|}
   A_0 & A_1 & (\neg A_0) & ((\neg A_0) \to A_1) & (\neg ( (\neg A_0) \to A_1)) \\
   \hline
   T   & F   & F          & T                    & F
   \end{array}$$

- $A_0 \lor A_1$

$$\begin{array}{|c|c|c|c|c|}
   A_0 & A_1 & (\neg A_0) & ((\neg A_0) \to A_1) & (A_0 \lor A_1) \\
   \hline
   T   & F   & F          & T                    & T
   \end{array}$$

- $A_0 \land A_1$

$$\begin{array}{|c|c|c|c|c|}
   A_0 & A_1 & (\neg A_1) & (A_0 \to (\neg A_1)) & (\neg ( A_0 \to (\neg A_1))) & (A_0 \land A_1) \\
   \hline
   T   & F   & T          & T                    & F                            & F
   \end{array}$$

The following solution of the problem asserts that only the truth values of the atomic sentences in the formula matter.

__Problem 2__ _Suppose $\delta$ is any formula and $u$ and $v$ are truth assignments such that $u(A_n) = v(A_n)$ for all atomic formulas $A_n$ which occur in $\delta$.  Then $u(\delta) = v(\delta)$._

- Proof by strong induction where $n$ is the number of connectives in a formula of $\mathcal{L}_P$

    + _Base step_ $(n = 0)$: If $\delta$ is a formula with 0 connectives it must be an atomic formula. Therefore $u(\delta) = v(\delta)$
    
    + _Induction hypothesis_ $(n \leq k)$: If $\delta$ is a formula with $n \leq k$ connectives then if $u$ and $v$ are truth assignments such that $u(A_n) = v(A_n)$ for all atomic formulas $A_n$ which occur in $\delta$ we have that $u(\delta) = v(\delta)$
    
    + _Induction step_ $(n = k + 1)$: If $\delta$ is a formula with $n = k + 1$ connectives then we have two possibilities:
    
        + $\delta = (\neg \alpha)$ where $\alpha$ has $n = k$ connectives
        
        + $\delta = (\alpha \to \beta)$ where $\alpha$ and $\beta$ have $n \leq k$ connectives
        
            + If $\delta = (\neg \alpha)$ then by the _induction hypothesis_ $u(\alpha) = v(\alpha)$. Therefore $u((\neg \alpha)) = v((\neg \alpha))$ so $u(\delta) = v(\delta)$
            
            + If $\gamma = (\alpha \to \beta)$ then by the _induction hypothesis_ $u(\alpha) = v(\alpha)$ and $u(\beta) = v(\beta)$. Therefore $u((\alpha \to \beta)) = v((\alpha \to \beta))$ so $u(\delta) = v(\delta)$
            
__Problem 3__ Suppose $u$ and $v$ are truth assignments such that $u(A_n) = v(A_n)$ for every atomic formula $A_n$.  Then $u = v$.

- $u$ and $v$ are functions therefore to prove that $u = v$ we have to check that [See [@mendelson_number_2008, Chapter 1, Section 1.18, Exercise 3]]:

    + $u$ and $v$ have the same domain
    
    + For any $x$ in the domain, $u(x) = v(x)$

      + The domain of $u$ and $v$ are the same and corresponds to all the formulas of $\mathcal{L}_P$

      + In __Problem 2__ we prove that if $u(A_n) = v(A_n)$ for all atomic formulas $A_n$ which occur in $\delta$ then $u(\delta) = v(\delta)$. Therefore for any $\delta$ that is a formula, $u(\delta) = v(\delta)$
      
- Taking into account the above results we have that $u = v$

__Problem 4__ _If $\alpha$ and $\beta$ are formulas and $v$ is a truth assignment,  then:_

- $v(\neg \alpha) = T$ if and only if $v(\alpha) = F$

    + If $v(\neg \alpha) = T$ by definition $v(\alpha) = F$
    
    + If $v(\alpha) = F$ assume that $v(\neg \alpha) = F$
    
        + If $v(\neg \alpha) = F$ by definition $v(\alpha) = T$. But $v(\alpha) = F$ and $v(\alpha) = T$ which is a contradiction because $v$ is a truth assigment. Therefore $v(\neg \alpha) = T$     

- $v(\alpha \to \beta) = T$ if and only if $v(\beta) = T$ whenever $v(\alpha) = T$

    + If $v(\alpha \to \beta) = T$ and $v(\alpha) = T$ by definition $v(\beta) = T$
    
    + If $v(\beta) = T$ and $v(\alpha) = T$ assume $v(\alpha \to \beta) = F$
    
        + If $v(\alpha \to \beta) = F$ then $v(\alpha) = T$ and $v(\beta) = F$. But $v(\beta) = T$ and $v(\beta) = F$ which is a contradiction because $v$ is a truth assigment. Therefore $v(\alpha \to \beta) = T$    


- $v(\alpha \land \beta) = T$ if and only if $v(\alpha) = T$ and $v(\beta) = T$

    + $\begin{split}
       v(\alpha \land \beta) = T & \iff v((\neg (\alpha \to (\neg \beta)))) = T \\
       & \iff v((\alpha \to (\neg \beta))) = F \\
       & \implies v(\alpha) = T \text{ and } v((\neg \beta)) = F \\
       & \iff v(\alpha) = T \text{ and } v(\beta) = T
       \end{split}$
       
    + If $v(\alpha) = T$ and $v(\beta) = T$ assume that $v(\alpha \land \beta) = F$.
    
        + If $v(\alpha \land \beta) = F$ then $v((\neg (\alpha \to (\neg \beta)))) = F$ so $v((\alpha \to (\neg \beta))) = T$. Hence $v(\alpha) = F$ or $v((\neg \beta)) = T$ which means that $v(\alpha) = F$ or $v(\beta) = F$. But we have also that $v(\alpha) = T$ and $v(\beta) = T$ which is a contradiction because $v$ is a truth assigment. Therefore $v(\alpha \land \beta) = T$      

- $v(\alpha \lor \beta) = T$ if and only if $v(\alpha) = T$ or $v(\beta) = T$

    + $\begin{split}
       v(\alpha \lor \beta) = T & \iff v(((\neg \alpha) \to \beta)) = T \\
       & \implies v((\neg \alpha)) = F \text{ or } v(\beta) = T \\
       & \iff v(\alpha) = T \text{ or } v(\beta) = T \\
       \end{split}$

    + If $v(\alpha) = T$ or $v(\beta) = T$ assume that $v(\alpha \lor \beta) = F$.
    
        + If $v(\alpha \lor \beta) = F$ then $v(((\neg \alpha) \to \beta)) = F$ so $v((\neg \alpha)) = T$ and $v(\beta) = F$. Hence $v(\alpha) = F$ and $v(\beta) = F$. But we have also that $v(\alpha) = T$ or $v(\beta) = T$ which is a contradiction because $v$ is a truth assigment.  Therefore $v(\alpha \lor \beta) = T$   

- $v(\alpha \leftrightarrow \beta) = T$ if and only if $v(\alpha) = v(\beta)$

    + $\begin{split}
       v(\alpha \leftrightarrow \beta) = T & \iff v(((\alpha \to \beta) \land (\beta \to \alpha))) = T \\
       & \iff v((\alpha \to \beta)) = T \text{ and } v((\beta \to \alpha)) = T \\
       & \implies v(\alpha) = F \text{ or } v(\beta) = T \text{ and } v(\beta) = F \text{ or } v(\alpha) = T \\
       & \iff v(\alpha) = F \text{ and } v(\beta) = F  \text{ or } v(\alpha) = T \text{ and } v(\beta) = T \\
       & \iff v(\alpha) = v(\beta)
       \end{split}$
       
    + If $v(\alpha) = v(\beta)$ assume $v(\alpha \leftrightarrow \beta) = F$
    
        + $\begin{split}
           v(\alpha \leftrightarrow \beta) = F & \iff v(((\alpha \to \beta) \land (\beta \to \alpha))) = F \\
           & \iff v((\alpha \to \beta)) = F \text{ or } v((\beta \to \alpha)) = F \\
           & \implies v(\alpha) = T \text{ and } v(\beta) = F \text{ or } v(\beta) = T \text{ and } v(\alpha) = F \\
           & \iff v(\alpha) \neq v(\beta) \\
           & \implies \text{There is a contradiction because } v(\alpha) = v(\beta) 
           \end{split}$
           
        + Therefore $v(\alpha \leftrightarrow \beta) = T$
        
__Definition__ If $v$ is a truth assignment and $\varphi$ is a formula, we will often say that $v$ _satisfies_ $\varphi$ if $v(\varphi) = T$.  Similarly, if $\Sigma$ is a set of formulas, we will often say that $v$ satisfies $\Sigma$ if $v(\sigma) = T$ for every $\sigma \in \Sigma$. We will say that $\varphi$ (respectively,  $\Sigma$) is _satisfiable_ if there is some truth assignment which satisfies it.

- $v \text{ satisfies } \varphi \iff v(\varphi) = T$

- $v \text{ satisfies } \Sigma \iff (\text{for all } \sigma)(\sigma \in \Sigma \implies v(\sigma) = T)$

__Definition__ A formula $\varphi$ is a _tautology_ if it is satisfied by every truth assignment.  A formula $\psi$ is a _contradiction_ if there is no truth assignment which satisfies it.

- $\varphi \text{ is a tautology} \iff (\text{for all } v)(v \text{ is a truth assigment} \implies v(\varphi) = T)$

- $\psi \text{ is a contradiction} \iff (\text{for all } v)(v \text{ is a truth assigment} \implies v(\psi) = F)$

__Problem 5__ _Prove that if $\alpha$ is any formula,  then $((\neg \alpha) \lor \alpha)$ is a tautology and $((\neg \alpha) \land \alpha)$ is a  contradiction._

$$\begin{array}{|c|c|c|c|c|}
   \text{Truth assigment} & \alpha & (\neg \alpha) & ((\neg \alpha) \lor \alpha) & ((\neg \alpha) \land \alpha) \\
   \hline
   v_1                    & T      & F             & T                           & F                              \\
   \hline
   v_2                    & F      & T             & T                           & F                              \\ 
   \end{array}$$

__Problem 6__ _Prove that a formula $\beta$ is a tautology if and only if $(\neg \beta)$ is a contradiction._

- $\begin{split}
   \beta \text{ is a tautology} & \iff (\text{for all } v)(v \text{ is a truth assigment} \implies v(\beta) = T) \\
   & \iff  (\text{for all } v)(v \text{ is a truth assigment} \implies v((\neg \beta)) = F) \\
   & \iff (\neg \beta) \text{ is a contradiction}
   \end{split}$
   
__Definition__ A set of formulas $\Sigma$ _implies_ a formula $\varphi$,  written as $\Sigma \vDash \varphi$,  if every truth assignment $v$ which satisfies $\Sigma$ also satisfies $\varphi$.  We will often write $\Sigma \nvDash \varphi$ if it is not the case that $\Sigma \models \varphi$.  In the case where $\Sigma$ is empty,  we will usually write $\vDash \varphi$ instead of $\emptyset \vDash \varphi$. Similarly,  if $\Delta$ and $\Gamma$ are sets of formulas,  then $\Delta$ _implies_ $\Gamma$,  written as $\Delta \vDash \Gamma$,  if every truth assignment $v$ which satisfies $\Delta$ also satisfies $\Gamma$.

- $\Sigma \vDash \varphi \iff (\text{for all } v)(v \text{ is a truth assigment and }(\text{for all } \sigma)(\sigma \in \Sigma \implies v(\sigma) = T) \implies v(\varphi) = T)$

- $\Sigma \nvDash \varphi \iff (\text{exists } v)(v \text{ is a truth assigment and }(\text{for all } \sigma)(\sigma \in \Sigma \implies v(\sigma) = T) \text{ and } v(\varphi) = F)$

    + Note that a formula $\varphi$ is a tautology if and only if $\vDash \varphi$,  and a contradiction if and only if $\vDash (\lnot \varphi)$
    
        + $\begin{split}
           \vDash \varphi & \iff \emptyset \vDash \varphi \\
           & \iff (\text{for all } v)(v \text{ is a truth assigment and }(\text{for all } \sigma)(\sigma \in \emptyset \implies v(\sigma) = T) \implies v(\varphi) = T) \\
           & \iff (\text{for all } v)(v \text{ is a truth assigment} \implies v(\varphi) = T) \\
           & \iff \varphi \text{ is a tautology}
           \end{split}$
           
        + $\begin{split}
           \vDash (\neg \varphi) & \iff \emptyset \vDash (\neg \varphi) \\
           & \iff (\text{for all } v)(v \text{ is a truth assigment and }(\text{for all } \sigma)(\sigma \in \emptyset \implies v(\sigma) = T) \implies v((\neg \varphi)) = T) \\
           & \iff (\text{for all } v)(v \text{ is a truth assigment and }(\text{for all } \sigma)(\sigma \in \emptyset \implies v(\sigma) = T) \implies v(\varphi) = F) \\
           & \iff (\text{for all } v)(v \text{ is a truth assigment} \implies v(\varphi) = F) \\
           & \iff \varphi \text{ is a contradiction}
           \end{split}$
           
            + Remember that $(\text{for all } \sigma)(\sigma \in \emptyset \implies v(\sigma) = T)$ is always true because $\sigma \in \emptyset$ is false

__Problem 7__ _If $\Gamma$ and $\Sigma$ are sets of formulas such that $\Gamma \subseteq \Sigma$,  then $\Sigma \vDash \Gamma$._

- $\begin{split}
    \sigma \in \Sigma \text{ and } v(\sigma) = T  \text{ and } \varphi \in \Gamma & \implies \varphi \in \Sigma \text{ and } v(\varphi) = T \\
    & \implies \Sigma \vDash \Gamma
    \end{split}$
    
__Problem 8__ _How can one check whether or not $\Sigma \vDash \varphi$ for a formula $\varphi$ and a finite set of formulas $\Sigma$?_

# Bibliography

