---
title: "Truth Assignments"
date: 2020-05-19 20:29:29
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 2
bibliography: a_problem_course_in_mathematical_logic.bib
link-citations: yes
---

<style>

body {
text-align: justify}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo      = TRUE, 
                      warning   = FALSE, 
                      message   = FALSE,
                      fig.align = "center")
```

```{r libraries}


```

# Truth Assignments

This section is based on [@bilaniuk_problem_2009, Chapter 2]

A formula $\varphi$ of $\mathcal{L}_P$ is true or false depending on how we interpret the atomic formulas which appear in $\varphi$.

- Since we don't want to commit ourselves to a single interpretation we will define how any assignment of _truth values_ $T$ ("true") and $F$ ("false") to atomic formulas of $\mathcal{L}_P$ can be extended to all other formulas.

We will also get a reasonable definition of what it means for a formula of $\mathcal{L}_P$ to follow logically from other formulas. 

__Definition__ A _truth assignment_ is a function $v$ whose domain is the set of all formulas of $\mathcal{L}_P$ and whose range is the set $\{ T, F \}$ of truth values,  such that:

- $v(A_n)$ is defined for every atomic formula $A_n$

- For any formula $\alpha$

    + $v((\neg \alpha)) = \begin{cases}
                           T & \text{if } v(\alpha) = F \\
                           F & \text{if } v(\alpha) = T
                           \end{cases}$

- For any formulas $\alpha$ and $\beta$

    + $v((\alpha \to \beta)) = \begin{cases}
                                F & \text{if } v(\alpha) = T \text{ and } v(\beta) = F \\
                                T & \text{Otherwise}
                                \end{cases}$
                                
Note that we have not defined how to handle any truth values besides $T$ and $F$ in $\mathcal{L}_P$. According to [@bilaniuk_problem_2009, Chapter 2] logics with other truth values have uses,  but are not relevant in most of mathematics.


A convenient way to write out the determination of the truth value of a formula on a given truth assignment is to use a _truth table_: 

- List all the subformulas of the given formula across the top in order of length

- Fill in their truth values on the bottom from left to right

    + Except for the atomic formulas at the extreme left,  the truth value of each subformula will depend on the truth values of the subformulas to its left
    
        + For example if $v(A_0) = T$ and $v(A_1) = F$ then to know the truth assignment of $v(((\neg A_1) \to (A_0 \to A_1)))$ we can use the following truth table:
        
$$\begin{array}{|c|c|c|c|c|}
   A_0 & A_1 & (\neg A_1) & (A_0 \to A_1) & ((\neg A_1) \to (A_0 \to A_1)) \\
   \hline
   T   & F   & T          & F             & F
   \end{array}$$

## Problems 1-4 and solutions

__Problem 1__ Suppose $v$ is a truth assignment such that $v(A_0) = v(A_2) = T$ and $v(A_1) = v(A_3) = F$.  Find $v(\alpha)$ if $\alpha$ is:

- $\neg A_2 \to \neg A_3$

- $\neg A_2 \to A_3$

- $\lnot ( \neg A_0 \to A_1)$

- $A_0 \lor A_1$

- $A_0 \land A_1$

# Bibliography

